# Clustering {#clustering}

Tutorial:  https://satijalab.org/seurat/articles/pbmc3k_tutorial#cluster-the-cells 


Clustering the cells will allow you to visualise the variability of your data, can help to segregate cells into cell types.   

Creating cell clusters the is easy, but understanding what is in them biologically-speaking can take time. We will try some things that aren't in the introductory seruat tutorial.



## Cluster cells

<!-- Do clustering -->
<!-- Choose a resolution -->
<!-- What are the clusters? -->
<!-- Find cluster markers -->
<!-- Name some. -->

<!-- Not a cfull challange, but encourage people to chhose a random cluster resolution and share --> 


<!-- #### Challenge: Try different cluster settings {- .challenge} -->

<!-- Run `FindNeighbours` and `FindClusters` again, with a different number of dimensions or with a different resolution. Examine the resulting clusters using `DimPlot`. -->

<!-- To maintain the flow of this tutorial, please put the output of this exploration in a different variable, such as `pbmc2`! -->

<!-- ####  {-} -->
<!-- ** -->


## Choosing a cluster resolution


Its a good idea to try different resolutions when clustering to identify the variability of your data. 

```{r}
# Do clustering at 0.1, 0.2, 0.3... 2.0
pbmc <- FindClusters(object = pbmc, reduction = "umap", resolution = seq(0.1, 2, 0.1), dims = 1:10)
# the different clustering created
names(pbmc@meta.data)

# How many clusters (and how many cells in those clusters) do we get at different resolutions?
table(pbmc$RNA_snn_res.0.1)
table(pbmc$RNA_snn_res.0.5)
table(pbmc$RNA_snn_res.2)
```

Plot a 'clustree' to decide how many clusters you have and what resolution capture them.

```{r fig.height=12}
library(clustree)
clustree(pbmc, prefix = "RNA_snn_res.") + theme(legend.key.size = unit(0.05, "cm"))
```

Name cells with the corresponding cluster name at the resolution you pick. This case we are happy with 0.5.
```{r}
# The name of the cluster is prefixed with 'RNA_snn_res' and the number of the resolution
Idents(pbmc) <- pbmc$RNA_snn_res.0.5
```

Plot the UMAP with coloured clusters with Dimplot

```{r}
DimPlot(pbmc, label = TRUE, repel = TRUE, label.box = TRUE) + NoLegend()
```
